\documentclass[a4paper,12pt]{article}

\usepackage{amsmath,amsthm,amssymb,ae,aecompl,sgame,natbib,sgamevar}
\usepackage[margin=2.5cm]{geometry}
\usepackage{tikz}
\usetikzlibrary{trees}

\title{Exercises Adv. Micro II}
\author{Christoph Schottm\"uller}

\begin{document}

\maketitle

\section{Rationalizability}
\begin{enumerate}
\item %Come up with a three player game in which each player has two actions and the actions surviving iterative elimination of strictly dominated strategies is not the same as the set of rationalizable actions.
  Show that in the following three player game (only payoffs of player 1 are given), $M$ is a never best response although $M$ is not strictly dominated.
\begin{table}[h]
    \centering
    \begin{tabular}{l|c|c}
      & L &R\\ \hline
      U& 10,  &10, \\
      M&6,  &0, \\
      D& 0, & 10,
    \end{tabular}
    $\qquad$
    \begin{tabular}{l|c|c}
      & L &R\\ \hline
      U& 10,  &0,  \\
      M&0,&6, \\
      D& 10, & 10,
    \end{tabular}
    \caption{P1 chooses row, P2 column, P3 table}
    \label{tab:domVsRatio}
  \end{table}
  
\item Show that the order of elimination does not matter for the outcome of iterative elimination of strictly dominated strategies. \\(Hint: Consider two orders, say A and B, and suppose they lead to different outcomes. Consider then the first action to be deleted in one procedure, say A, while surviving the other, i.e. B. Show that this action is strictly dominated at the end of procedure B which yields the desired contradiction.)
\item (Numerical) Consider two players participating in a lab experiment. The monetary payoffs the two players get from the experimenter when choosing their strategies are given in table \ref{tab:monRatio}. (note: this time the numbers in the table are not utilities but amounts of money!) Assume that players are only interested in their own monetary payoffs and not in the monetary payoffs for the other players.
  \begin{table}[h]
    \centering
    \begin{tabular}{l|c|c}
      & L &R\\ \hline
      U& 3,1  &0,0  \\
      M1&1,0&1,1 \\
      M&2.6,0  &0.5,0 \\
      D& 0,1 & 3,0
    \end{tabular}
    \caption{Monetary payoff}
    \label{tab:monRatio}
  \end{table}
  \begin{enumerate}
  \item Suppose players value money linearly $u_i(x)=x$ for any monetary amount $x$. Which actions are rationalizable?
  \item Suppose players value money  $u_i(x)=-e^{-r x}$ for any monetary amount $x$. Which actions are rationalizable if the parameter $r$ is very high? (you can choose a high number like $3$ and check)
     \item Suppose players value money $u_i(x)=e^{r x}$ for any monetary amount $x$. Which actions are rationalizable if the parameter $r$ is very high?
     \end{enumerate}
     What is the interpretation in terms of risk preferences? (see \cite{weinstein2016effect} for more on this topic)
\end{enumerate}


\section{Existence of Nash equilibrium}
\label{sec:nash-theorem-fixed}

\begin{enumerate}
\item Show that a pure strategy Nash equilibrium exists in the following two player game: $I=\{P1,P2\}$, $S_1=S_2=[0,1]$, $u_i(s_i,s_{-i})$ is differentiable and strictly (quasi-)concave in $s_i$ and $\partial u_i/\partial s_i$ is continuous in both variables.\\
  (Hint: Show that the first order condition leads to a continuous best response function. Then, verify that the two best response functions must intersect.)
\end{enumerate}

\section{Correlated equilibrium}
\label{sec:corr-equil}
\begin{enumerate}
\item Find a correlated equilibrium that yields a higher payoff to every player than the unique Nash equilibrium (game in table \ref{tab:corEq}).
   \begin{table}[h]
\centering
 \begin{game}{3}{3}
       \> L \> C \>R\\ %\hline
U   \>1,1   \> 2,4 \> 4,2   \\
M   \>4,2   \> 1,1 \> 2,4 \\
D  \> 2,4    \>4,2  \>1,1
\end{game}  
\caption{correlated equilibrium}
\label{tab:corEq}
\end{table}
\item In the example from the lecture (using in particular the inequalities on slide 9), show that the set of correlated equilibria is the set of all probability distributions $(p_{UL},p_{UR},p_{DL},p_{DR})$ for which the the probability on each pure Nash equilibrium is at least as high as the probability on any other strategy profile. (Note: This is \emph{not} a general result!) Given this, derive the correlated equilibrium with the lowest and the highest sum of payoffs. 
\end{enumerate}

\section{Bayesian Nash equilibrium}
\label{sec:Bayes-nash-eq}
\begin{enumerate}
\item Two player Cournot with uncertain costs: Payoffs of player $i$ are $(\theta _i-q_1-q_2)q_i$ where $\theta _i$ is the difference between the intercept of a linear demand function and costs. It is commonly known that $\theta _1=1$. However, $\theta _2$ is private information of firm 2. Firm 1's (commonly known) belief is that $\theta _2$ is $3/4$ with probability $1/2$ and $5/4$ with probability $1/2$.
  \begin{enumerate}
  \item If firm 1's equilibrium quantity is $q_1^*$, what is firm 2's best response?
  \item If type $j\in\{l,h\}$ of firm 2 produces quantity $q_2^j$, what is firm 1's best response quantity? (i.e. maximizing expected payoff)
    \item What is the BNE of this game?
  \end{enumerate}
\item Purification: Consider the game in table \ref{tab:purify}. Let $\theta _i$ be private information and be distributed uniformly on $[0 ,\varepsilon ]$. Assuming $\varepsilon >0$ sufficiently small, derive a BNE in cutoff strategies; i.e. P1 (P2) plays U (L) if $\theta _1>\bar{\theta }_1$ ($\theta _2>\bar{\theta }_2$) and D (R) otherwise for some $\bar \theta_i\in(0,\bar{\varepsilon })$. Compare the equilibrium in the limit as $\varepsilon \rightarrow 0$ to the mixed strategy equilibrium in the game without private information (i.e. the game where $\theta _i=0$ is commonly known). 
     \begin{table}[h]
\centering
 \begin{game}{2}{2}
       \> L  \>R\\ %\hline
U   \>5$+\theta_1$,1$+\theta _2$    \> $\theta_1$,0   \\
D  \> 4,4$+\theta _2$    \>1,5
\end{game}  
\caption{purification}
\label{tab:purify}
\end{table}
\end{enumerate}

\section{Auctions}
\label{sec:auctions}

\begin{enumerate}

\item Take the first price sealed bid auction in the independnet private value framework. The purpose of this exercise is to show that in every BNE, the bidding strategy of an arbitrary bidder $i$ is weakly increasing in his valuation. Take two values $v_i^h$ and $v_i^l$ with $v_i^h>v_i^l$ and let $s_i()$ be an equilbrium strategy of player $i$.
  \begin{enumerate}
  \item In equilibrium type $v_i^l$ must prefer bidding $s_i(v_i^l)$ to bidding $s_i(v_i^h)$. Write down the corresponding inequality.
    % (v_i^l-s_i(v_i^l))*prob(win,s_i(v_i^l))\geq (v_i^l-s_i(v_i^h))*prob(win,s_i(v_i^h))
  \item In equilibrium type $v_i^h$ must prefer bidding $s_i(v_i^h)$ to bidding $s_i(v_i^l)$. Write down the corresponding inequality.
    % (v_i^h-s_i(v_i^h))*prob(win,s_i(v_i^h))\geq (v_i^h-s_i(v_i^l))*prob(win,s_i(v_i^l))
  \item Use the two inequality to show that $s_i(v_i^h)\geq s_i(v_i^l)$. (Hint: Note that bidding higher increases the probability of getting the object in a first price auction!)
    % adding inequalities gives: $(v_i^h-v_i^l) (prob(win,s_i(v_i^h))-prob(win,s_i(v_i^l))\geq 0$ which is only true if $s_i(v_i^h)\geq s_i(v_i^l)$
  \end{enumerate}
  
% \item Two people have to decide who gets an indivisible item. Their valuations of the item are private information and we regard the two $v_i$ as drawn independently from a uniform distribution on $[0,1]$. The two people have come up with the following auction mechanism: They simultaneously state a bid $b_i$ and the one stating the higher bid gets the item but has to pay half of his bid to the other player.
%   \\
%   Assume that there is a symmetric equilibrium of this game in which players bid according to a strictly increasing strategy $s(v)$. (If you find this hard, assume that $s(v)=\alpha v$ for some $\alpha>0$ and determine below the $\alpha$ for which this is an equilibrium.)
%   \begin{enumerate}
%   \item Write down the expected payoff of a player bidding $b$ (given that the other player bids according to $s$). (Hint: Do not forget that he might receive a payment from the other player!)
%     % $\int_0^{t(b)}v_i-b/2 \,dv_j+\int_{t(b)}^1 s(v_j)/2\,dv_j$
%   \item Derive the first order condition of an optimal bid.
%     % $t'(b)(v_i-b/2)-t(b)/2-t'(b) b/2=0$
%   \item What is the equilibrium bidding strategy? Why is it not optimal to bid your true valuation?
%     % use $t(b)=v_i$, $b=s(v_i)$ and $t'(b)=1/s'(v_i)$ in foc to get $(v_i-s(v_i))/(s'(v_i))-v/2=0$ which is solved by $s(v_i)=v_i*2/3$
%   \end{enumerate}
\item A buyer's valuation is uniformly distributed on $[0,1]$ and the buyer's private information. A seller's cost is also uniformly distributed on $[0,1]$ and the seller's private information. The buyer states an ask price $b_b$ and the seller simultaneously states an offer price $b_s$. If $b_s<b_b$, trade takes place at price $(b_s+b_b)/2$ while otherwise trade does not take place.
  \begin{enumerate}
  \item Explain/show why $b_s(c)=c$ and $b_b(v)=v$ is not an equilibrium although it would maximize welfare. 
  \item Show that the following is an equilibrium for any given $p\in[0,1]$:
    \begin{equation*}
      b_b(v)=
      \begin{cases}
        p & \text{ if }v\geq p\\
        0 & \text{ else }
      \end{cases}
    \qquad
      b_s(c)=
      \begin{cases}
        p & \text{ if }c\leq p\\
        1 & \text{ else }.
      \end{cases}
    \end{equation*}
    \item Suppose that players use linear strategies $b_b(v)=\alpha+\beta v$ and $b_s(c)=\gamma+\delta c$ with $\beta,\delta>0$. Taking the strategy of the other player derive the best response of a seller/buyer with valuation $v$ / cost $c$. For which values of $\alpha,\beta,\gamma,\delta$ form the linear strategies an equilibrium?
    \end{enumerate}
  \item Two players participate in a contest: Each player chooses an effort level $e_i$. Effort is costly and reduces player $i$'s payoff by  $\theta _ie_i$. The player who choose the highest effort wins a prize of value 1. That is, the winner has payoff $1-\theta _i e_i$ while the loser has payoff $-\theta _i e_i$. The $\theta _i$ are private information and each player views the other player's type as being uniformly distributed on $[0.1,1.1]$. \\
    Write player $i$'s maximization problem given that the other player uses a strictly decreasing strategy $s$ (with inverse $t$). Derive the first order condition for optimal effort choice. Solve for a symmetric equilibrium in strictly decreasing strategies.
    % $\max_{e}(1.1-t(e))-\theta _i e$ where $t$ is the inverse of the other player's strategy $s$. Foc: $-t'(e)-\theta _i=0$, in equilibrium $t'(e(\theta _i))=1/s'(\theta _i)$ and we get $s'(\theta _i)=-1/\theta _i$. Using $s(1.1)=0$, we get $s(\theta _i)=\int_{\theta _i}^{1.1}1/x \,dx=log(1.1)-log(\theta _i)$
    (More on such contests can be found in \cite{moldovanu2001optimal}.)
\end{enumerate}

\section{Subgame perfect Nash equilibrium}
\label{sec:subgame-perfect-nash}

\begin{enumerate}
\item $I$ Players play the same strategic form game $\Gamma=[I,\{S_i\},\{u_i\}]$ $T$ times. Players observe immediately after each round $t=1,\dots, T$ which actions were chosen by the other players. Assume that $\Gamma$ has a unique Nash equilibrium $\sigma$. Show that the extensive form game $\Gamma_E$ formed by the $T$ time repetition of $\Gamma$ has a unique subgame perfect Nash equilibrium. (extra exercise: find an example $\Gamma$ with a unique NE in which $\Gamma_E$ has several NE that are not subgame perfect.)
  % add a "total destruction" action yielding for both players -1000 to a normal prisoner's dilemma and let $T=2$; the strategy cooperate in 1; defect in 2 if cooperation in 1 and total destruction otherwise is a NE but not subgame perfect
\end{enumerate}

\section{Weak PBE and sequential equilibrium}
\label{sec:weak-pbe-sequential}

\begin{enumerate}
\item Find a pure strategy NE in the game of figure \ref{fig:selten_ex}. Show that this NE is subgame perfect but not sequentially rational. Find a weak PBE. (Hint: use mixed strategies)
   \begin{figure}[h]
\centering
% First, set the overall layout of the tree
% You might need to play with these sizes to ensure nothing overlaps.
\tikzstyle{level 1}=[level distance=1.5cm, sibling distance=6cm]
\tikzstyle{level 2}=[level distance=1.5cm, sibling distance=2.5cm]
\tikzstyle{level 3}=[level distance=1.5cm, sibling distance=2cm]
\tikzstyle{level 4}=[level distance=1.5cm, sibling distance=1.5cm]
\begin{tikzpicture}
%Start with the parent node, and slowly build out the tree
% with each "child" representing a new level of the diagram
% each "node" represents a labelled (or unlabeled if you 
% want) node in the diagram.
\node{1}
    child{
             node{2}
             child{
               node{0,0,0}
             edge from parent
             node[left]{$L_2$}}
             child{
               node(b){3}
                  child{
               node{4,4,0}
               edge from parent
               node[left]{l}
               }
             child{
               node{0,2,1}
               edge from parent
               node[right]{r}
               }
                edge from parent
                node[right]{$R_2$}
               }
           edge from parent
           node[left,above]{$L_1$}
           }
    child{
           child{
               node(c){3}
                  child{
               node{1,1,1}
               edge from parent
               node[left]{l}
               }
             child{
               node{1,1,0}
               edge from parent
               node[right]{r}
               }
               }
         edge from parent
         node[right,above]{$R_1$}
         };
\draw [dashed](c)--(b);
\end{tikzpicture}
\caption{Selten's horse}
\label{fig:selten_ex}
\end{figure}

\item Consider the two games in figures \ref{fig:seq_eq_coalescing} and \ref{fig:seq_eq_coalescing2}. To what extent are these games different or the same? Show that there is a pure strategy sequential equilibria in figure \ref{fig:seq_eq_coalescing} that does not exist in \ref{fig:seq_eq_coalescing2}. Would this also be true for  weak PBE instead of sequential equilibrium?
% (R,r) supported by belief (0,1) is seq eq in first game but there is no equivalent in the second as A strictly dominates B and therefore seq rationality requires P1 to play A instead of B in the subgame. For any completely mixed strat converging to ?A P2's belief must be (1,0) and therefore his action in seq eq is l which means (LA,l) is unique seq eq in second game. Note that (RA,r) with belief (0,1) is still a weak PBE as P2's info set in never reached.
  
   \begin{figure}[h]
\centering
% First, set the overall layout of the tree
% You might need to play with these sizes to ensure nothing overlaps.
\tikzstyle{level 1}=[level distance=1.5cm, sibling distance=3.5cm]
\tikzstyle{level 2}=[level distance=1.5cm, sibling distance=1.5cm]
\tikzstyle{level 3}=[level distance=1.5cm, sibling distance=2cm]
\tikzstyle{level 4}=[level distance=1.5cm, sibling distance=1.5cm]
\begin{tikzpicture}
%Start with the parent node, and slowly build out the tree
% with each "child" representing a new level of the diagram
% each "node" represents a labelled (or unlabeled if you 
% want) node in the diagram.
\node{1}
             child{
               node(a){2}
                  child{
               node{5,1}
               edge from parent
               node[left]{l}
               }
             child{
               node{1,0}
               edge from parent
               node[right]{r}
               }
               edge from parent
               node[left,above]{A}
               }
             child{
               node(b){2}
                  child{
               node{0,0}
               edge from parent
               node[left]{l}
               }
             child{
               node{0,1}
               edge from parent
               node[right]{r}
               }
               edge from parent
               node[right]{B}
               }
    child{
         node{3,3}
         edge from parent
         node[right, above]{R}
         };
\draw [dashed](a)--(b);
\end{tikzpicture}
\caption{sequential equilibrium and coalescing moves}
\label{fig:seq_eq_coalescing}
\end{figure}

 \begin{figure}[h]
\centering
% First, set the overall layout of the tree
% You might need to play with these sizes to ensure nothing overlaps.
\tikzstyle{level 1}=[level distance=1.5cm, sibling distance=3.5cm]
\tikzstyle{level 2}=[level distance=1.5cm, sibling distance=3.5cm]
\tikzstyle{level 3}=[level distance=1.5cm, sibling distance=2cm]
\tikzstyle{level 4}=[level distance=1.5cm, sibling distance=1.5cm]
\begin{tikzpicture}
%Start with the parent node, and slowly build out the tree
% with each "child" representing a new level of the diagram
% each "node" represents a labelled (or unlabeled if you 
% want) node in the diagram.
\node{1}
    child{
             node{1}
             child{
               node(a){2}
                  child{
               node{5,1}
               edge from parent
               node[left]{l}
               }
             child{
               node{1,0}
               edge from parent
               node[right]{r}
               }
               edge from parent
               node[left]{A}
               }
             child{
               node(b){2}
                  child{
               node{0,0}
               edge from parent
               node[left]{l}
               }
             child{
               node{0,1}
               edge from parent
               node[right]{r}
               }
               edge from parent
               node[right]{B}
               }
           edge from parent
           node[left]{L}
           }
    child{
         node{3,3}
         edge from parent
         node[right]{R}
         };
\draw [dashed](a)--(b);
\end{tikzpicture}
\caption{game above with additional move}
\label{fig:seq_eq_coalescing2}
\end{figure}

\end{enumerate}

\section{Signaling}
\label{sec:signaling}

\begin{enumerate}
% \item Take the jobmarket model from the slides. Suppose that the worker has an outside option, i.e. if he is offered only wages of less than $r>0$, than he will refuse all offers and take his outside option (of value $r$). The value of the outside option does not depend on type. What are the equilibria if $r\in(\theta_l,\mathbb{E}[\theta]$? What are the equilibria if $r>\mathbb{E}[\theta ]$? How do you evaluate the welfare effect of the possibility to signal?
\item Consider the least cost separating equilibrium of the jobmarket model on the slides. Suppose the government introduces a tax on high wages (all wages above $\mathbb{E}[\theta ]$) and a subsidy for low wages (all others). The tax is $t$ and the subsidy is $t*\lambda/(1-\lambda)$, i.e. the tax subsidy scheme has a balanced budget. Show that this (i) tax scheme leads to a Pareto improvement if $\lambda$ is sufficiently large (and $t$ is chosen not too large) and (ii) high productivity types are worse off as a result of the tax if $\lambda$ is low. 
  % Firms always get zero profits and $\theta_l$ types clearly are made better off by the subsidy.  A marginal increase in $t$ from $t=0$ improves welfare of high types if their cost of effort is reduced by more than the tax burden. The marginal effect of the tax on payoff is $-1$. The reduction in effort (at $t=0$) is determined from the indifference condition $\theta_h-t-c(e,\theta _l)=\theta _l+t\lambda/(1-\lambda)$ via implicit function theorem, i.e. $d\tilde e/dt=-1/((1-\lambda)*c_e(\tilde e,\theta_l))$. Hence, welfare of $\theta_j$ goes up by $\frac{1}{1-\lambda}\frac{c_e(\tilde e,\theta_h)}{c_e(\tilde{e},\theta _l)}$. For $\lambda$ sufficiently close to 1 this is higher than 1, for $\lambda$ sufficiently close to zero it is less than 1 by the single crossing assumption. 
\item Exercise 13.C.6 in MWG.
\end{enumerate}

\section{Cheap talk}
\label{sec:cheap-talk}

\begin{enumerate}
\item The state of the world $\theta $ is either 0 or 1. Both, 0 and 1, have probability 1/2. An expert observes the state and then sends a cheap talk message to a decision maker. The available messages are ``high'' (h) and ``low'' (l). (Assume in the following that the messages are used in accordance with their name, i.e. the high message leads to a higher belief of the decision maker.) The decision maker chooses an action $a\in [0,2]$ with the objective of maximizing $(a-\theta )^2$. The objective of the expert is $-(a-\theta-b)^2$ for some $b\in(0,1)$.
  \begin{enumerate}
  \item What is a mixed strategy for the expert? % probability of sending high message for each state
    \item Intuitively, which state will the expert reveal trughtfully? % high state as expert always wants higher action than DM
    \item Given a mixed strategy of the expert, what are the decision maker's beliefs when receiving a low/high message? % in state 0, let prob of h be q (while in state 1 it is 1), then $mu(l)=0$ and $\mu(h)=\frac{.5*1}{.5*1+0.5*q}=1/(1+q)$
      \item Given a mixed strategy of the expert, what is the best response of the decision maker? % $min \mu * (a-1)^2+(1-\mu)(a)^2$ which gives foc: $a=\mu$ 
      \item What is the equilibrium strategy of the expert? %truthtelling is eq if $-b^2\geq-(1-b)^2$, i.e. if $b\leq 1/2$; otherwise indifference of expert in 0 state: $b^2=(1/(1-q)-b)^2$ solved for $q=1-2b$
      \end{enumerate}
    \item Take the model from the lecture but assume that the state is distributed not uniformly but with some arbitrary distribution $\Phi$ (with strictly positive density $\phi$) on $[0,1]$. Take a partitional equilibrium as in the lecture. Show that $a(m)=\mathbb{E}[\theta |m]$. Show then that S benefits from commitment in expectation, i.e. if S could commit to always truthfully reveal the state he would in expectation be better off than in equilibrium (where the expectation is taken over the state of the world $\theta $).\footnote{We use here the assumption that the utility is a quadratic loss function. I think the result can be generalized to the case $u^R(a,\theta )=u(a-\theta)$ and $u^S(a,\theta )=u(a-\theta -b)$ if $u$ is a strictly concave function. However, already for strictly quasiconcave $u$ functions the result is no longer true. }
      % S's exp utility in each partition element $\int -(E[theta|m]-\theta-b)^2 d\Phi(\theta|m) = -var(\theta|m)-b^2+0>-b^2$
    \item (Extra exercise) A patient can be in one of three health states: A, B, C. State A and B have prior probability 2/5 and C has 1/5. Doctor and patient both receive a private signal about the health state that is either 0 or 1. The following table describes the signal distribution.
      \begin{table}[h]
        \centering
        \begin{tabular}{l|c|c|c}
          $\sigma$&A&B&C\\ \hline
          (0,0) & 0 &0&1\\
          (0,1) & 0&4/5&0\\
          (1,0)&1/5&1/5&0\\
          (1,1)&4/5&0&0
        \end{tabular}
        \caption{Signal}
        \label{tab:signcifd}
      \end{table}

The interpretation is that, given health state A, signal $(\sigma^p,\sigma^d)=(1,1)$ occurs with probability 4/5 where $\sigma^p$ is the patient's signal etc.

There are three treatments: a,b,c. The patient's payoff as function of state and treatment and the costs of each treatment are given in the following table. (There is some interpretation that $a$ is effective but expensive, $c$ is no treatment; A is severe problem, B is slight problem and C is healthy; under- and overtreatment both reduce the patient's payoff.) 
\begin{table}[h]
  \centering
  \begin{tabular}{l|c|c|c|c}
    &A&B&C&costs\\ \hline
    a&8&9.7&9.2&5\\
    b&4&9&9.6&3\\
    c&0&5&10&1
  \end{tabular}
  \caption{Payoffs and costs}
  \label{tab:payoffcifd}
   \begin{tabular}{l|c|c|c}
    &A&B&C\\ \hline
    a&3&4.7&4.2\\
    b&1&6&6.6\\
    c&-1&4&9
  \end{tabular}
  \caption{welfare: payoff - costs}
  \label{tab:welfarecifd}
\end{table}

The timing of the game is that the patient can send a cheap talk message in $\{0,1\}$ about his signal to the doctor. The doctor then determines the treatment.

\begin{enumerate}
\item Let welfare be patient payoff minus costs. What are the welfare maximizing treatments in each state? What are the treatments maximizing patient payoff (i.e. neglecting the costs)? 
\item Assume doctor and patient have the objective of maximizing the patient's payoff (and ignore costs because the patient is insured). What is the equilibrium? (If you see that there are more than 1 equilibrium, use the one you find most reasonable for the following questions.)
\item Assume that the patient's objective is to maximize his payoff while the doctor maximizes welfare. Is there an equilibrium in which the patient truthfully reveals his signal to the doctor? (If not, assume in the following that there is a babbling equilibrium in which the doctor relies only on his own signal.)
\item Is welfare higher when the doctor maximizes welfare or when the doctor maximizes patient payoff?
\item Show that costs are higher when the doctor maximizes welfare if the prior is $(1/2-p_c/2,1/2-p_c/2,p_c)$ for $p_c>0$ small enough.
  \item What implications do these results have for designing the health care system?
\end{enumerate}

This exercise will not be covered in class. For a solution, see \cite{schottmueller2013cifd}, in particular section 3.

\end{enumerate}

\section{Revelation principle}
\label{sec:revelation-principle}

\begin{enumerate}
\item A policy $x\in[0,1]$ has to be selected (e.g. a tax rate). There are $I\geq 3$ agents and each agent $i$ has a most preferred policy $\theta _i$ called $i$'s ``peak''. The payoff of agent $i$ if policy $x$ is selected is $-|x-\theta _i|$. We consider two social choice functions:
  \begin{itemize}
  \item mean selection: $f(\theta _1,\dots,\theta _I)=\sum_{i=1}^I\theta _i/I$
    \item median selection: select the median of $\{x_1,\dots,x_I\}$
  \end{itemize}
  \begin{enumerate}
  \item Let each $\theta _i$ be uniformly and independently be distributed on $[0,1]$. Is mean selection implementable by any mechanism $\Gamma$? Is median selection implementable by any mechanism $\Gamma$?
    \item Find a distribution $\Phi$ such that both policies are implementable if each $\theta _i$ is independently distributed according to $\Phi$.
    \end{enumerate}
    \item MWG 23.D.1
    \end{enumerate}

\section{Dominant strategy implementation}
\label{sec:domin-strat-impl}

\begin{enumerate}
\item VCG with positive costs: A public project costs $c=3$. There are three agents with valuations $\theta _1=2$, $\theta _2=0$ and $\theta _3=2$. Calculate the transfers in the Pivot mechanism and determine the players payoffs. Is there anything potentially problematic?
\item There is a seller with privately known costs $c$ and a buyer with privately known valuation $v$. Both, $v$ and $c$ are (from the other players' point of view) uniformly distributed on $[0,1]$. The seller has payoff $t_S-c x$ where $t_S$ is a monetary payment the seller receives and $x$ the probability of trade. The buyer has payoff $vx-t_B$.
  \begin{enumerate}
  \item What is the efficient decision?
  \item What are the Pivot payments?
    \item Can you spot something problematic with the Pivot payments?
    \end{enumerate}
%seller is paid $v$ and buyer pays only $c$ if trade takes place (0 otherwise)
\item (Extra exercise: probably not covered in class) A country with $I$ citizens has to decide how much it wants to reduce emissions. Denote the fraction of the current level by which emissions should be reduced as $x\in[0,1]$. Each citizen has a private valuation of emission reduction which is $u_i(x)=\theta _ix-t_i$ (where $\theta _i$ is private information and $t_i$ is a tax/transfer). Reducing emissions by $x$ will reduce production and therefore wealth by $x^2$. Assume that the support of the distribution of $\theta _i$ is $[0,2/I)$.
  \begin{enumerate}
  \item What is the efficient level of emission reduction $x$?
    \item Calculate the Pivot mechanism transfers for this problem.
  \end{enumerate}
\end{enumerate}

\bibliographystyle{chicago}
\bibliography{/home/christoph/stuff/bibliography/references.bib}

\end{document}
